\begin{thebibliography}{99}
  \bibitem{QM} D.\ J.\ Griffiths: {\em Introduction to Quantum Mechanics}, Pearson Prentice Hall (2005).

  \bibitem{ES} R.\ M.\ Martin: {\em Electronic Structure}, Cambridge University Press (2004).
  
  \bibitem{KS} W.\ Kohn and L.\ J.\ Sham: {\em Self-Consistent Equations Including Exchange and Correlation Effects}, Physical Review A {\bf 140}, 1133 (1965).

  \bibitem{LDA} J.\ M.\ MacLaren, D.\ P.\ Clougherty, M.\ E.\ McHenry and M.\ M.\ Donovan: {\em Parameterised local spin density exchange-correlation energies and potentials for electronic structure calculations}, Computer Physics Communications {\bf 66}, 383--391 (1991).

  \bibitem{NIST} S.\ Kotochigova, Z.\ H.\ Levine, E.\ L.\ Shirley, M.\ D.\ Stiles and C.\ W.\ Clark: {\em Atomic Reference Data for Electronic Structure Calculations}, Physical Measurement Laboratory, NIST.\\ Link: \url{http://www.nist.gov/pml/data/dftdata/}
  
  \bibitem{WB} M.\ Weissbluth: {\em Atoms and Molecules}, Academic Press (1978).
  
  \bibitem{EM} D.\ J.\ Griffiths: {\em Introduction to Electrodynamics}, Prentice Hall (1999).
  
  \bibitem{CP} N.\ J.\ Giordano and H.\ Nakanishi: {\em Computational Physics}, Pearson Prentice Hall (2006).
  
  \bibitem{NR} W.\,H.\,Press, S.\,A.\,Teukolsky, W.\,T.\,Vetterling and B.\,P.\,Flannery: {\em Numerical Recipes in C}, Cambridge University Press (1992).

  \bibitem{MC} S.\ E.\ Koonin and D.\ C.\ Meredith: {\em Compututational Physics}, Westview Press (1990).
  
  \bibitem{RI} G.\ Racah: {\em Theory of Complex Spectra.~I}, Physical Review {\bf 61}, 186 (1942).

  \bibitem{RII} G.\ Racah: {\em Theory of Complex Spectra.~II}, Physical Review {\bf 62}, 438 (1942).
  
  \bibitem{RIII} G.\ Racah: {\em Theory of Complex Spectra.~III}, Physical Review {\bf 63}, 367 (1943).
  
  \bibitem{RIV} G.\ Racah: {\em Theory of Complex Spectra.~IV}, Physical Review {\bf 76}, 1352 (1949).
  
  \bibitem{YBK} E.\ Koch: {\em Exchange Mechanisms}, in \cite{YB}.
  
  \bibitem{GBK} E.\ Koch: {\em Many-Electron States}, in \cite{GB}.
  
  \bibitem{YBE} R.\ Eder: {\em Multiplets in Transition Metal Ions}, in \cite{YB}.
  
  \bibitem{YB} E.\ Pavarini, E.\ Koch, F.\ Anders and M.\ Jarrell (eds.): {\em Correlated Electrons: From Models to Materials}, Reihe Modeling and Simulation, Vol.\ 2, Forschungszentrum J\"{u}lich (2012).
  
  \bibitem{GB} E.\ Pavarini, E.\ Koch and U.\ Schollw\"{o}ck (eds.): {\em Emergent Phenomena in Correlated Matter}, Reihe Modeling and Simulation, Vol.\ 3, Forschungszentrum J\"{u}lich (2013).
\end{thebibliography}
